---
title: "cqThemes: Ces Quarto Themes - Consistent Styling for Statistical Reports and Presentations"
author: "Cesaire Tobias"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cosmo
    highlight: tango
vignette: >
  %\VignetteIndexEntry{cqThemes: Consistent Styling for Statistical Reports and Presentations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  fig.align = "center",
  dpi = 120
)
```

# Introduction

The `cqThemes` package provides consistent styling for statistical reports and presentations created with Quarto or R Markdown. It includes themes for three different output formats:

1. **PDF Reports** - Professional styling for academic and business reports
2. **HTML Documents** - Interactive web-based documents with responsive design
3. **RevealJS Presentations** - Slide decks with consistent branding

Each theme comes with coordinated color palettes, typography, and spacing to ensure a professional and consistent look across all your documents.

## Installation

You can install the package from GitHub:

```{r eval=FALSE}
# Install devtools if you don't have it
# install.packages("devtools")

# Install cqThemes from GitHub
devtools::install_github("ces0491/cqThemes")
```

# PDF Report Styling

For PDF reports, the package provides a clean, professional theme with typography and spacing optimized for readability.

## Basic Usage

To apply the PDF theme to your Quarto or R Markdown document, simply load the package and call the `use_pdf_theme()` function in your setup chunk:

```{r eval=FALSE}
library(cqThemes)
use_pdf_theme()
```

This sets up all the necessary styling for your document.

## Plot Styling

The `theme_cq_pdf()` function provides a consistent theme for ggplot2 visualizations in PDF documents:

```{r}
library(cqThemes)
library(ggplot2)

# Create a sample dataset
set.seed(9104)
df <- data.frame(
  x = rnorm(100),
  y = rnorm(100),
  group = factor(sample(1:3, 100, replace = TRUE))
)

# Create a plot with the PDF theme
ggplot(df, aes(x = x, y = y, color = group)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "Sample Scatterplot with PDF Theme",
    subtitle = "Demonstrating cqThemes styling",
    x = "X Variable",
    y = "Y Variable"
  ) +
  theme_cq_pdf() +
  scale_color_manual(values = cq_colors(3, type = "categorical"))
```

## Table Formatting

For tables in PDF documents, use the `format_table_pdf()` function:

```{r}
# Load example data
data(mtcars)

# Format a table for PDF output
format_table_pdf(
  head(mtcars[, 1:5]),
  caption = "Sample Table with PDF Formatting",
  digits = 2,
  highlight_row = 3,
  font_size = 10
)
```

# HTML Document Styling

For HTML documents, the package provides a responsive theme with interactive elements.

## Basic Usage

To apply the HTML theme to your document, call the `use_html_theme()` function:

```{r eval=FALSE}
library(cqThemes)
use_html_theme(interactive = TRUE)
```

This sets up all the necessary styling and enables interactive plots.

## Interactive Plots

The HTML theme automatically converts ggplot2 visualizations to interactive plotly charts:

```{r}
# Create a plot with the HTML theme
p <- ggplot(df, aes(x = x, y = y, color = group)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "Interactive Scatterplot with HTML Theme",
    subtitle = "Hover over points for more information",
    x = "X Variable",
    y = "Y Variable"
  ) +
  theme_cq_html() +
  scale_color_manual(values = cq_colors(3, type = "categorical"))

# In an actual HTML document, this would be converted to an interactive plotly chart
# Here we'll just show the static version
p
```

## Interactive Tables

For interactive tables in HTML documents, use the `format_table_html()` function:

```{r}
# Load example data
data(mtcars)

# Format a table for HTML output
if (requireNamespace("DT", quietly = TRUE)) {
  format_table_html(
    head(mtcars),
    caption = "Sample Interactive Table",
    digits = 2,
    buttons = c("copy", "csv"),
    compact = TRUE
  )
} else {
  knitr::kable(head(mtcars), caption = "Sample Table (DT package not available)")
}
```

# RevealJS Presentation Styling

For presentations, the package provides a clean, minimalist theme with larger text and simplified elements.

## Basic Usage

To apply the RevealJS theme to your presentation, call the `use_revealjs_theme()` function:

```{r eval=FALSE}
library(cqThemes)
use_revealjs_theme()
```

## Presentation Plots

The `theme_cq_revealjs()` function provides a theme optimized for presentations:

```{r}
# Create a plot with the RevealJS theme
ggplot(df, aes(x = x, y = y, color = group)) +
  geom_point(size = 4, alpha = 0.8) +
  labs(
    title = "Presentation Plot with RevealJS Theme",
    x = "X Variable",
    y = "Y Variable"
  ) +
  theme_cq_revealjs() +
  scale_color_manual(values = cq_colors(3, type = "categorical"))
```

## Presentation Tables

For tables in presentations, use the `format_table_reveal()` function:

```{r}
# Format a table for RevealJS presentations
format_table_reveal(
  head(mtcars[, 1:5], 5),
  caption = "Sample Table for Presentations",
  digits = 1,
  highlight_row = 2
)
```

# Color Palettes

The package provides several color palette functions to ensure consistent coloring across all visualizations.

## Default Colors

The `cq_colors()` function provides the default color palette:

```{r}
# Display the default color palette
par(mar = c(2, 1, 2, 1))
barplot(
  rep(1, 10),
  col = cq_colors(10, type = "categorical"),
  main = "Default Categorical Color Palette",
  horiz = TRUE,
  axes = FALSE
)
text(0.5, 1:10 - 0.5, paste("Color", 1:10), pos = 4)
```

## Sequential and Diverging Palettes

For ordered data, use the sequential and diverging palettes:

```{r}
# Display sequential color palettes
par(mfrow = c(2, 2), mar = c(2, 1, 2, 1))

# Blue sequential
barplot(
  rep(1, 7),
  col = cq_sequential_palette(7, palette = "blue"),
  main = "Blue Sequential",
  horiz = TRUE,
  axes = FALSE
)

# Green sequential
barplot(
  rep(1, 7),
  col = cq_sequential_palette(7, palette = "green"),
  main = "Green Sequential",
  horiz = TRUE,
  axes = FALSE
)

# Teal-Brown diverging
barplot(
  rep(1, 7),
  col = cq_diverging_palette(7, palette = "teal-brown"),
  main = "Teal-Brown Diverging",
  horiz = TRUE,
  axes = FALSE
)

# Blue-Red diverging
barplot(
  rep(1, 7),
  col = cq_diverging_palette(7, palette = "blue-red"),
  main = "Blue-Red Diverging",
  horiz = TRUE,
  axes = FALSE
)
```

# Additional Utilities

The package includes several utility functions to help with common tasks.

## Correlation Plots

The `create_correlation_plot()` function creates consistently styled correlation plots:

```{r}
# Create a correlation matrix
cor_matrix <- cor(mtcars[, 1:5])

# Create a correlation plot
create_correlation_plot(
  cor_matrix,
  title = "Correlation Matrix of Car Features",
  colors = colorRampPalette(cq_colors(2, type = "main"))(200)
)
```

## Saving Plots

Use the `save_plot()` function to save plots with consistent settings:

```{r eval=FALSE}
# Create a plot
p <- ggplot(mtcars, aes(x = mpg, y = hp)) +
  geom_point() +
  labs(title = "Horsepower vs MPG") +
  theme_cq_pdf()

# Save the plot
save_plot(p, "my_plot.png", width = 8, height = 6, dpi = 300)
```

# Templates

The package includes document templates for all three output formats. You can access these templates from RStudio:

1. File → New File → R Markdown... → From Template
2. Select one of the cqThemes templates:
   - cqThemes PDF Report
   - cqThemes HTML Document
   - cqThemes RevealJS Presentation

# FAQ and Troubleshooting

## Common Issues

### Missing Packages

If you see an error about missing packages, use the `ensure_packages()` function:

```{r eval=FALSE}
cqThemes::ensure_packages(c("ggplot2", "plotly", "DT", "kableExtra"))
```

### Theme Not Applied

Make sure you're calling the appropriate theme function at the beginning of your document:

```{r eval=FALSE}
# For PDF:
use_pdf_theme()

# For HTML:
use_html_theme()

# For RevealJS:
use_revealjs_theme()
```

## Contact

For bug reports or feature requests, please file an issue on GitHub: [https://github.com/username/cqThemes/issues](https://github.com/username/cqThemes/issues)

# Conclusion

The `cqThemes` package provides a consistent styling system for statistical reports and presentations. By using the provided themes, color palettes, and formatting functions, you can create professional, visually appealing documents with minimal effort.
